---
title: "Getting Started"
---

```{r, include = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE, out.width = '100%')
```


```{r, setup}
library(sdwisard)
library(dplyr)
library(ggplot2)
```

Make sure the package is installed. See [installation instructions here](https://caopenwater.github.io/sdwisard/).  

## Basic Query

The **PSID** (public water system ID number) of a water system is the only mandatory argument to `get_data()`. Use the PSID of a system to get all the water quality data for that system. 

For example, Alameda County Water District is PSID 0110001:  

```{r}
alameda_all <- get_data(psid = "0110001")
glimpse(alameda_all)
```

## Helper Functions

**sdwisard** includes three helper functions:

* `get_water_system()` - Find PSID by county name 
* `get_analyte_summary()` - Summary about available data for a water system  
* `get_storet_id()` - Find storet id with common chemical name 

Supply a county to `get_water_system()` to return PSIDs and water system names within that county.  
```{r}
get_water_system(county = "Alameda")
```

Supply a PSID to `get_analyte_summary()` to view a summary of a water system's available analyte data. 
```{r}
get_analyte_summary(psid = "0110001") 
```

Use the function `get_storet_id` to search for storet IDs by common names of an analytes.

```{r}
nitrate_ids <- get_storet_id("nitrate")
nitrate_ids
```

Provide the storet ID to query for a specific analyte. The following query return all nitrate data (storet ID = A-029) for Alameda county Watere District. 
```{r}
alameda_nitrate <- get_data(psid = "0110001", storet = "A-029")
```

We can plot these results with `ggplot` as follows:

```{r}
alameda_nitrate %>% 
  mutate(date = as.Date(SAMP_DATE)) %>% 
  ggplot(aes(date, FINDING)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "NITRATE + NITRITE (AS N) at Alameda County Water District",
       y = "(MG/L)") +
  theme_minimal()
```

To query a specific date range provide a `start_date` and/or `end_date` argument(s) in **YYYY-MM-DD** format.  

```{r}
alameda_09_2018 <- get_data(psid = "0110001", start_date = "2018-09-01", end_date = "2018-09-30")
```


## Built-in package objects

Full tables of metadata are also available:  

```{r}
water_systems
```

```{r}
psid_analyte
```

```{r}
analytes
```

